
#1. Preliminaries
## a. Install and load the following R packages:
```{r}
library(bnlearn)
library(pcalg)
library(igraph)
library(Rgraphviz)

data("insurance")

#colnames(insurance)

#summary(insurance)

```

## b.Create the insurance ground truth model from the model string

```{r}
modelstring = paste0("[Age][Mileage][SocioEcon|Age][GoodStudent|Age:SocioEcon]",
                     "[RiskAversion|Age:SocioEcon][OtherCar|SocioEcon][VehicleYear|SocioEcon:RiskAversion]",
                     "[MakeModel|SocioEcon:RiskAversion][SeniorTrain|Age:RiskAversion]",
                     "[HomeBase|SocioEcon:RiskAversion][AntiTheft|SocioEcon:RiskAversion]",
                     "[RuggedAuto|VehicleYear:MakeModel][Antilock|VehicleYear:MakeModel]",
                     "[DrivingSkill|Age:SeniorTrain][CarValue|VehicleYear:MakeModel:Mileage]",
                     "[Airbag|VehicleYear:MakeModel][DrivQuality|RiskAversion:DrivingSkill]",
                     "[Theft|CarValue:HomeBase:AntiTheft][Cushioning|RuggedAuto:Airbag]",
                     "[DrivHist|RiskAversion:DrivingSkill][Accident|DrivQuality:Mileage:Antilock]",
                     "[ThisCarDam|RuggedAuto:Accident][OtherCarCost|RuggedAuto:Accident]",
                     "[MedCost|Age:Accident:Cushioning][ILiCost|Accident]",
                     "[ThisCarCost|ThisCarDam:Theft:CarValue][PropCost|ThisCarCost:OtherCarCost]")

g_truth = model2network(modelstring)
#plot(dag)

```

##c check ground truth class 
```{r}

class(g_truth)

```

## d.get adjacency matrix from
```{r}
adj_matrix=amat(g_truth)

#adj_matrix
```

## e.Build a directed igraph network from the adjacency matrix
```{r}
network_insurance=g=graph_from_adjacency_matrix(adj_matrix,mode="directed",weighted = T, diag=F  )
E(network_insurance)$arrow.size <- .4
plot(network_insurance,layout=layout_with_dh,
     edge.width=abs(E(network_insurance)$weight),
     edge.color= "blue")
legend(x=-0.8,y=-1.5,legend = "Ground Truth insurance network")
```

#2. Hill Climbing

## b.Reconstruct the insurance network using the hill-climbing approach
```{r}
res_hc=hc(insurance)
#plot(res_hc)
class(res_hc)#HC function returns un object of class bn
```

## c.Get the adjacency matrix
```{r}
adj_mat_hc=amat(res_hc)
```
## d.Build a directed igraph network from the adjacency matrix

```{r}

net_hc=igraph::graph_from_adjacency_matrix(adj_mat_hc,mode="directed",weighted = T, diag = F)
E(net_hc)$arrow.size <- .4

par(mfrow = c(1,2))
plot(network_insurance,layout=layout_with_dh,
     edge.width=abs(E(network_insurance)$weight),
     edge.color= "blue")
legend(x=-1.2,y=-1.5,legend = "Ground Truth insurance network")

plot(net_hc,layout=layout_with_gem,
     edge.width=abs(E(net_hc)$weight),
     edge.color= "red")
legend(x=-1.2,y=-1.5,legend = "HC Reconstructed graph")

```
## e.Count the number of true positive (TP), false positive (FP) and false negative (FN)

```{r}

res_comp=bnlearn::compare(bnlearn::skeleton(g_truth),bnlearn::skeleton(res_hc))
res_comp

precision=res_comp$tp/(res_comp$tp+res_comp$fp)
print(paste0('precision= ',precision))

recall=res_comp$tp/(res_comp$tp+res_comp$fn)
print(paste0('recall= ',recall))

Fscore=(2*precision*recall)/(precision+recall)
print(paste0('Fscore= ',Fscore))

```

##f  
```{r}
graphviz.compare(res_hc,g_truth, diff.args = list(fp.col = "red", fp.lwd = 4, show.first = FALSE))
                 

```
# PC method
## a.Reconstruct the insurance network using the PC approach

```{r}
dt=data.matrix(insurance)
dt=dt-1

level_names=c()
for(i in colnames(dt)){
  a=nlevels(as.factor(dt[,i]))
  level_names=c(a,level_names)
}
suffStat<-list(dm=dt,nlev=level_names,adaptDF=FALSE)
pc.D<-pc(suffStat,indepTest=disCItest,alpha=0.01,labels=colnames(dt),verbose = F)
```

## b.Get the adjacency matrix

```{r}
adj_matrix_pc=amat(as.bn(pc.D))
```

### Build a directed igraph network from the adjacency matrix 

```{r}

g3=graph_from_adjacency_matrix(adj_matrix_pc,mode="undirected")

```

### c.to plot a nice graph

```{r}
g_pc <- bnlearn::as.graphNEL(as.bn(pc.D))

par(mfrow = c(1,2))
plot(g_pc,  attrs=list(node = list(fillcolor = "yellow", fontsize=80)),main="Nice printed Dag")
plot(g3,  attrs=list(node = list(fillcolor = "green", fontsize=40)),main="Original Dag")
```

## d.Count the number of true positive (TP), false positive (FP) and false negative (FN)

```{r}
res = bnlearn::compare(bnlearn::skeleton(as.bn(g_pc)),bnlearn::skeleton(g_truth) )

prec = res$tp/(res$tp + res$fp)
print(paste0('precision= ',prec))

recal = res$tp/(res$tp + res$fn)
print(paste0('recall= ',recal))

fscore = (2*prec*recal)/(prec+recal)
print(paste0('Fscore= ',fscore))

```

## e.Highlight the edges

```{r}

graphviz.compare(g_truth, as.bn(g_pc),diff.args = list(fp.col = "red", fp.lwd = 4, show.first = FALSE))

```

#Aracne method
## a.Reconstruct the insurance network using the PC approach
```{r}
dt_aracne=as.data.frame(insurance)

aracne_graph=bnlearn::aracne(dt_aracne)
```

## b.Get the adjacency matrix

```{r}
adj_matrix4 =amat(aracne_graph)
```

## c.Build a directed igraph network from the adjacency matrix

```{r}
g4=graph_from_adjacency_matrix(adj_matrix4,mode="undirected")

g_aracne <- bnlearn::as.graphNEL(aracne_graph) # use this to avoid printing of graphviz.plot

par(mfrow = c(1,2))
plot(g4,  attrs=list(node = list(fillcolor = "orange", fontsize=30)),main="Dag")
plot(g_aracne,  attrs=list(node = list(fillcolor = "lightgreen", fontsize=60)),main="Dag")
```

## d.Count the number of true positive (TP), false positive (FP) and false negative (FN)

```{r}
res_aracne = bnlearn::compare(bnlearn::skeleton(as.bn(g_aracne)),bnlearn::skeleton(g_truth) )

prec_aracne = res_aracne$tp/(res_aracne$tp + res_aracne$fp)
prec_aracne
recal_aracne = res_aracne$tp/(res_aracne$tp + res_aracne$fn)
recal_aracne
fscore_aracne = (2*prec_aracne*recal_aracne)/(prec_aracne+recal_aracne)
fscore_aracne
```

## e.Highlight the edges

```{r}
graphviz.compare(g_truth, as.bn(g_aracne),diff.args = list(fp.col = "red", fp.lwd = 4, show.first = FALSE))

```

